---
title: "kaya"
---

### The Kaya Identity

The Kaya Identity is a simple model for observing Earth's anthropogenic system macroscopically.[^1]

[^1]: This is a re-working of Jean-Marc Jancovinci's excellent article

$$
F=  P \cdot \frac {G}{P} \cdot \frac{E}{G} \cdot \frac {F}{E}
$$

Where:

-   *F* = CO~2~ emissions from human sources
-   *P* = global population
-   *G* = Gross Domestic Produc
-   *E* = global energy consumption

And:

-   *G/P* = per capita GDP

-   *E/G* is the [energy intensity](https://en.wikipedia.org/wiki/Energy_intensity "Energy intensity") of the [GDP](https://en.wikipedia.org/wiki/Gross_domestic_product "Gross domestic product")

-   *F/E* is the [emission intensity](https://en.wikipedia.org/wiki/Emission_intensity "Emission intensity") of energy

A mathematical identity is an equality where both sides of an equation are identical in value. It allows us to slowly peel the onion on one side of the equation.

This *identical* nature becomes evident as P, E, G will cancel themselves out, leaving us simple with:

$$
Wealth=  Population \cdot \frac {G.D.P}{Population} \cdot \frac{E}{G} 
$$

But, it becomes easier to read when the constituent parts are more explicit. Hence the following is identical:

$$
{\displaystyle CO_2=Population\cdot {\frac {GDP}{Population}}\cdot {\frac {EnergyUse}{GDP}}\cdot {\frac {CO_2}{EnergyUse}}}
$$

$$
CO_2 = Population \cdot\frac{CO_2}{Energy} \cdot\frac{Energy}{GDP}
\cdot\frac{GDP}{Population}
\cdot\frac{Population} {GDP}
$$

The first observation is that we live in a delicate balance with a burning ball of fire next to us. The surface of the earth is reflective, and sends the sun's energy back into space. But with amount of Carbon in the air prevents

$$
 CO_2 = CO_2
$$

$$
CO_2 = P\cdot\frac{CO_2}{E}
$$

$$
\text{ CO2 emissions = Carbon content of energy × Energy intensity of the economy × Production per person × Population}
$$

The largest Carbon Credits ETF is the KraneShares Global Carbon Strategy ETF KRBN with $557.92M in assets. In the last trailing year, the best-performing Carbon Credits ETF was KCCA at 12.52%. The most recent ETF launched in the Carbon Credits space was the Global X Carbon Credits Strategy ETF NTRL on 05/24/23.

```{r,, echo=FALSE}
# Loads tidyquant, lubridate, xts, quantmod, TTR 
library(tidyverse)
library(tidyquant)
library(plotly)
```

```{r}


tq_index_options()

getSymbols("CARB.TO",
             from = "2022-01-01",
             to = "2022-12-31")

stock <- data.frame(CARB.TO$CARB.TO.Adjusted)



stock <- data.frame(stock,rownames(stock))
colnames(stock) <- append('GOOG','date')

fig <- plot_ly(stock, type = 'scatter', mode = 'lines')%>%
  add_trace(x = ~date, y = ~GOOG, name = 'GOOG')%>%
  layout(showlegend = F)
options(warn = -1)

fig
```
